---
title: "Modeling Cold weather with Prophet"
date: 2022/02/09
output: html_document
---
  
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, fig.height = 5)
```


# Question: Will cold snaps in Boston be more rare?

*This is the model analysis of the project by [Kent J](https://github.com/kent37/Cold_winters)*
  
In this analysis, let's fit a model to predict [the cold weather](https://www.ncdc.noaa.gov) in Boston using {prophet} package for modeling sesonalities in time series. 
For further information and quick start with using prophet please see:
https://facebook.github.io/prophet/ or type `prophet::quick_start` in the help tab of your pane, a *quick_start.html* file will open up with some information.
  

```{r libs_mod, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(lubridate)
library(prophet)
```


```{r data_mod, message=FALSE, warning=FALSE, paged.print=FALSE}
path = 'Logan_weather_12-2012_3-2021.csv'
df = read_csv(path)
```


## Preprocessing data

In order to use Propphet we need a dataframe made of two columns:

- ds: date variable
- y: the dependent variable (or the outcome)


```{r temps_mod}
temps = df %>% 
  filter(REPORT_TYPE...3=='SOD') %>% 
  select(DATE, matches('Daily.*DryBulbTemperature')) %>% 
  mutate(date=ymd_hms(DATE),
         year=year(date)) %>% 
  select(year, date,
         min="DailyMinimumDryBulbTemperature",
         max="DailyMaximumDryBulbTemperature",
         avg="DailyAverageDryBulbTemperature")
temps<- temps%>% select(-year)
head(temps)
```

### Use prophet

Important is to have complete information about the date variable, and we can run 3 models:

- minimum temperature model
- maximun temperature model
- average temperature model

then combine them to see the overal the trend.

```{r df_mod}
min_df <- temps%>%select(ds=date,y=min)
max_df<- temps%>%select(ds=date,y=max)
avg_df <- temps%>%select(ds=date,y=avg)
```


#### Min model
```{r min_pro_mod, cache=TRUE}
min_mod <- prophet(min_df)
min_future <- prophet::make_future_dataframe(min_mod, periods=365)
min_forecast <- predict(min_mod,min_future)
```
#### Max model
```{r max_pro_mod, cache=TRUE}
max_mod <- prophet(max_df)
max_future <- prophet::make_future_dataframe(max_mod, periods=365)
max_forecast <- predict(max_mod,max_future)
```


#### Avg model
```{r avg_pro_mod, cache=TRUE}
avg_mod <- prophet(avg_df)
avg_future <- prophet::make_future_dataframe(avg_mod, periods=365)
avg_forecast <- predict(avg_mod,avg_future)
```

Let's see the outcome of our model with a first exploration using the `plot()` function.

```{r plot1_mod}
library(rafalib)
mypar(3)
plot(min_mod,min_forecast)
plot(max_mod,max_forecast)
plot(avg_mod,avg_forecast)
```

Then what we want is to combine the results of the three models, so we use {ggplot2} fro making the customization.

##### Make the plot with ggplot2
```{r plot2_mod}
  ggplot()+
  
  geom_point(data=min_df,mapping=aes(x=ds,y=y),size=0.2,color="red") + # original data
  geom_point(data=max_df,mapping=aes(x=ds,y=y),size=0.05,color="orange") + # original data
  geom_point(data=avg_df,mapping=aes(x=ds,y=y),size=0.05,color="midnightblue") + # original data
  
  geom_line(data=min_forecast,aes(ds,yhat),color="red",size=0.3) +
  geom_line(data=max_forecast,aes(x=ds,y=yhat_upper),size=0.05,color="darkred")+
  geom_line(data=avg_forecast,aes(x=ds,y=yhat_lower),size=0.05,color="darkred")
```

In particular the lower level of temperatures have this trend:
```{r plot3_mod}
ggplot()+
geom_line(data=min_df,aes(x=ds,y=y),size=0.3,color="grey55")+
geom_point(data=min_forecast,aes(x=ds,y=yhat),size=0.5) 
```

The function `prophet_plot_components()` represents the trend of the temperatures at different levels:

- overal trend
- day of the week
- day of the year

To see when the picks happen. Here are represented the trends for the minimum temperature model. But it can be replicated with the max and average values.
```{r plot4_mod}
prophet_plot_components(min_mod, min_forecast)
```


Finally, here we can see the main points where the temperatures pick to the lowest values, and identify when this happend.
```{r plot5_mod}
plot(min_mod, min_forecast) + add_changepoints_to_plot(min_mod)
```


Adjust the strength of the sparse prior using the input argument changepoint_prior_scale. Larger values allow the model to fit larger seasonal fluctuations, smaller values dampen the seasonality. 
By default, this parameter is set to 0.05. 

```{r plot6_mod}
min_mod2 <- prophet(min_df, changepoint.prior.scale = 1.5, daily.seasonality=TRUE)
forecast <- predict(min_mod2, min_future)
plot(min_mod2, min_forecast)
```

---
